---
title: "Main R Markdown Document For NCAA Basketball Analysis"
author: "Ben Wasserman and Connor Kennedy"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
library('magrittr')
library('caret')
library(plotROC)
```


## Table of Contents

### Document Prep and File Read In

### Munge Tournament Data

  - Parse ranking from seven different sources
  - Add rankings to the detailed results
  - Add team conferences to the detailed results
  - Randomly code each team for binomial models
  - Feature transformation to create various attributes
  
### Munge Regular Season Data

  - Join rankings with regular season data
  - Join with conferences
  - Randomly code teams for binomial models
  - Feature transformation to create various attributes
  - Get end of season averages for in-game statistics
  - Join with tournament data
  
### Add Viewers and Ratings

### Summary Statistics and EDA (This will have graphics too)

### Logistic Regression
  
  - In Sample
  - Out of Sample
  - With Regular Season In Game Data

### Ridge Model

### Lasso Model

### Random Forest Model

### Linear Regression on Viewer Data


Read in the files
```{r}
# remove environment variables
rm(list = ls())

# read in files
tourney_games <- read.csv("ConferenceTourneyGames.csv")
tourney_compact_results <- read.csv("NCAATourneyCompactResults.csv")
regular_season_compact_results <- read.csv("RegularSeasonCompactResults.csv")
teams <- read.csv("Teams.csv")
team_conferences <- read.csv("TeamConferences.csv")
tourney_seeds <- read.csv("NCAATourneySeeds.csv")
tourney_seeds_round_slots <- read.csv("NCAATourneySeedRoundSlots.csv")
tourney_detailed_result <- read.csv("NCAATourneyDetailedResults.csv")
tourney_slots <- read.csv("NCAATourneySlots.csv")
regular_detailed_result <- read.csv("RegularSeasonDetailedResults.csv")
rankings <- read.csv("MasseyOrdinals.csv")
rating_round_viewers <- read.csv("rating_round_viewers.csv")
tourney_semi_views_ratings <- read.csv("semi_rating_round_viewers.csv")

```

Parse 7 different rankings
```{r}
tourney_rankings <- rankings %>% filter(RankingDayNum > 132) %>% rename (DayNum = RankingDayNum) %>% select(
  Season, SystemName, TeamID, everything()
)

# get rid of duplicates by system name, season, and team ID
tourney_rankings <- tourney_rankings[!duplicated(rankings[1:3]),]

# parse rankings
bih_rankings <- tourney_rankings %>% filter(SystemName == 'BIH')
col_rankings <- tourney_rankings %>% filter(SystemName == 'COL')
dol_rankings <- tourney_rankings %>% filter(SystemName == 'DOL')
mor_rankings <- tourney_rankings %>% filter(SystemName == 'MOR')
wlk_rankings <- tourney_rankings %>% filter(SystemName == 'WLK')
wob_rankings <- tourney_rankings %>% filter(SystemName == 'WOB')
wol_rankings <- tourney_rankings %>% filter(SystemName == 'WOL')

# rename each the team IDs and the ranks to prepare for join
bih_rankings <- bih_rankings %>% rename(WTeamID = TeamID,
                                        WRankBih = OrdinalRank)

col_rankings <- col_rankings %>% rename(WTeamID = TeamID,
                                        WRankCol = OrdinalRank)

dol_rankings <- dol_rankings %>% rename(WTeamID = TeamID,
                                        WRankDol = OrdinalRank)


wlk_rankings <- wlk_rankings %>% rename(WTeamID = TeamID,
                                        WRankWlk = OrdinalRank)

wob_rankings <- wob_rankings %>% rename(WTeamID = TeamID,
                                        WRankWob = OrdinalRank)


wol_rankings <- wol_rankings %>% rename(WTeamID = TeamID,
                                        WRankWol = OrdinalRank)

mor_rankings <- mor_rankings %>% rename(WTeamID = TeamID,
                                        WRankMor = OrdinalRank)

# join with tourney detailed result
tourney_d_result <-left_join(x = tourney_detailed_result,
                                    y = bih_rankings,
                                    by = c("Season", "WTeamID"))


tourney_d_result <-left_join(x = tourney_d_result,
                                    y = wob_rankings,
                                    by = c("Season", "WTeamID"))


tourney_d_result <-left_join(x = tourney_d_result,
                                    y = wol_rankings,
                                    by = c("Season", "WTeamID"))


tourney_d_result <-left_join(x = tourney_d_result,
                                    y = wlk_rankings,
                                    by = c("Season", "WTeamID"))


tourney_d_result <-left_join(x = tourney_d_result,
                                    y = dol_rankings,
                                    by = c("Season", "WTeamID"))


tourney_d_result <-left_join(x = tourney_d_result,
                                    y = col_rankings,
                                    by = c("Season", "WTeamID"))


tourney_d_result <-left_join(x = tourney_d_result,
                                    y = mor_rankings,
                                    by = c("Season", "WTeamID"))


# rename rankings to prepare for join with losing team
bih_rankings <- bih_rankings %>% rename(LTeamID = WTeamID,
                                        LRankBih = WRankBih)

col_rankings <- col_rankings %>% rename(LTeamID = WTeamID,
                                        LRankCol = WRankCol)

dol_rankings <- dol_rankings %>% rename(LTeamID = WTeamID,
                                        LRankDol = WRankDol)

wlk_rankings <- wlk_rankings %>% rename(LTeamID = WTeamID,
                                        LRankWlk = WRankWlk)

wob_rankings <- wob_rankings %>% rename(LTeamID = WTeamID,
                                        LRankWob = WRankWob)

wol_rankings <- wol_rankings %>% rename(LTeamID = WTeamID,
                                        LRankWol = WRankWol)

mor_rankings <- mor_rankings %>% rename(LTeamID = WTeamID,
                                        LRankMor = WRankMor)


tourney_d_result <-left_join(x = tourney_d_result,
                                    y = bih_rankings,
                                    by = c("Season", "LTeamID"))

tourney_d_result <-left_join(x = tourney_d_result,
                                    y = wob_rankings,
                                    by = c("Season", "LTeamID"))

tourney_d_result <-left_join(x = tourney_d_result,
                                    y = wol_rankings,
                                    by = c("Season", "LTeamID"))

tourney_d_result <-left_join(x = tourney_d_result,
                                    y = wlk_rankings,
                                    by = c("Season", "LTeamID"))

tourney_d_result <-left_join(x = tourney_d_result,
                                    y = dol_rankings,
                                    by = c("Season", "LTeamID"))

tourney_d_result <-left_join(x = tourney_d_result,
                                    y = col_rankings,
                                    by = c("Season", "LTeamID"))

tourney_d_result <-left_join(x = tourney_d_result,
                                    y = mor_rankings,
                                    by = c("Season", "LTeamID"))

# remove unecessary columns
tourney_d_result = tourney_d_result[,-grep("^DayNum.",colnames(tourney_d_result))]
tourney_d_result = tourney_d_result[,-grep("^SystemName",colnames(tourney_d_result))]

```


Add rankings to the detailed results
```{r}
# just get the average ranking for each
tourney_d_result <- tourney_d_result %>% mutate(
  WAvgSeed = ((WRankMor + WRankWol + WRankBih + WRankWob + WRankWlk + WRankDol
               + WRankCol)/7),
  LAvgSeed = ((LRankMor + LRankWol + LRankBih + LRankWob + LRankWlk + LRankBih
               + LRankCol)/7),
  AvgSeedDiff = WAvgSeed - LAvgSeed
)

# get rid of tournament seeds prior to 2002 for consistency
tourney_seeds <- subset(tourney_seeds, Season > 2002)
tourney_seeds <- tourney_seeds %>% rename(WTeamID = TeamID)

tourney_d_result <-left_join(x = tourney_d_result,
                                    y = tourney_seeds,
                                    by = c("WTeamID", "Season"))

tourney_seeds <- tourney_seeds %>% rename(LTeamID = WTeamID)

tourney_d_result <-left_join(x = tourney_d_result,
                                    y = tourney_seeds,
                                    by = c("LTeamID", "Season"))

tourney_d_result <- tourney_d_result %>% rename(
  WTeamSeed = Seed.x,
  LTeamSeed = Seed.y
)

# make all the seeds only numeric
tourney_d_result$WTeamSeed <- as.integer(gsub('[a-zA-Z]', '',
                                                     tourney_d_result$WTeamSeed))
tourney_d_result$LTeamSeed <- as.integer(gsub('[a-zA-Z]', '',
                                                     tourney_d_result$LTeamSeed))

```

Add team conferences to detailed result
```{r}
team_conferences <- team_conferences %>% rename(WTeamID = TeamID)

# join with conferences
tourney_d_result <- left_join(x = tourney_d_result,
                                     y = team_conferences, 
                                     by = c("WTeamID", "Season"))

team_conferences <- team_conferences %>% rename(LTeamID = WTeamID)

tourney_d_result <- left_join(x = tourney_d_result,
                                     y = team_conferences, 
                                     by = c("LTeamID", "Season"))

tourney_d_result <- tourney_d_result %>% rename(
  WTeamConf = ConfAbbrev.x,
  LTeamConf = ConfAbbrev.y 
)

tourney_d_result <- tourney_d_result %>% mutate(LTeamConfFactor = fct_lump(LTeamConf, 10))

tourney_d_result <- tourney_d_result %>% mutate(WTeamConfFactor = fct_lump(WTeamConf, 10))


```


Randomly code each team to prepare for a binomial model
```{r}
# higher_team_won is the variable we will be predicting on
tourney_d_result <- tourney_d_result %>% mutate(
  lower_team = pmin(WTeamID, LTeamID),
  higher_team = pmax(WTeamID, LTeamID),
  higher_team_won = ifelse(higher_team == WTeamID, 1, 0),
  HTSeed = ifelse(higher_team == WTeamID, WTeamSeed, LTeamSeed),
  LTSeed = ifelse(lower_team == WTeamID, WTeamSeed, LTeamSeed),
  HTScore = ifelse(higher_team == WTeamID, WScore, LScore),
  LTScore = ifelse(lower_team == WTeamID, WScore, LScore),
  HTConf = ifelse(higher_team == WTeamID, as.character(WTeamConf), as.character(LTeamConf)),
  LTConf = ifelse(lower_team == WTeamID, as.character(WTeamConf), as.character(LTeamConf)),
  HTFGM = ifelse(higher_team == WTeamID, WFGM, LFGM),
  LTFGM = ifelse(lower_team == WTeamID, WFGM, LFGM),
  HTFGA = ifelse(higher_team == WTeamID, WFGA, LFGA),
  LTFGA = ifelse(lower_team == WTeamID, WFGA, LFGA),
  HTFGM3 = ifelse(higher_team == WTeamID, WFGM3, LFGM3),
  LTFGM3 = ifelse(lower_team == WTeamID, WFGM3, LFGM3),
  HTFGA3 = ifelse(higher_team == WTeamID, WFGA3, LFGA3),
  LTFGA3 = ifelse(lower_team == WTeamID, WFGA3, LFGA3),
  HTFTM = ifelse(higher_team == WTeamID, WFTM, LFTM),
  LTFTM = ifelse(lower_team == WTeamID, WFTM, LFTM),
  HTFTA = ifelse(higher_team == WTeamID, WFTA, LFTA),
  LTFTA = ifelse(lower_team == WTeamID, WFTA, LFTA),
  HTOR = ifelse(higher_team == WTeamID, WOR, LOR),
  LTOR = ifelse(lower_team == WTeamID, WOR, LOR),
  HTDR = ifelse(higher_team == WTeamID, WDR, LDR),
  LTDR = ifelse(lower_team == WTeamID, WDR, LDR),
  HTAst = ifelse(higher_team == WTeamID, WAst, LAst),
  LTAst = ifelse(lower_team == WTeamID, WAst, LAst),
  HTTO = ifelse(higher_team == WTeamID, WTO, LTO),
  LTTO = ifelse(lower_team == WTeamID, WTO, LTO),
  HTStl = ifelse(higher_team == WTeamID, WStl, LStl),
  LTStl = ifelse(lower_team == WTeamID, WStl, LStl),
  HTBlk = ifelse(higher_team == WTeamID, WBlk, LBlk),
  LTBlk = ifelse(lower_team == WTeamID, WBlk, LBlk),
  HTPF = ifelse(higher_team == WTeamID, WPF, LPF),
  LTPF = ifelse(lower_team == WTeamID, WPF, LPF),
  HTBih = ifelse(higher_team == WTeamID, WRankBih, LRankBih),
  LTBih = ifelse(lower_team == WTeamID, WRankBih, LRankBih),
  HTCol = ifelse(higher_team == WTeamID, WRankCol, LRankCol),
  LTCol = ifelse(lower_team == WTeamID, WRankCol, LRankCol),
  HTDol = ifelse(higher_team == WTeamID, WRankDol, LRankDol),
  LTDol = ifelse(lower_team == WTeamID, WRankDol, LRankDol),
  HTMor = ifelse(higher_team == WTeamID, WRankMor, LRankMor),
  LTMor = ifelse(lower_team == WTeamID, WRankMor, LRankMor),
  HTWlk = ifelse(higher_team == WTeamID, WRankWlk, LRankWlk),
  LTWlk = ifelse(lower_team == WTeamID, WRankWlk, LRankWlk),
  HTWob = ifelse(higher_team == WTeamID, WRankWob, LRankWob),
  LTWob = ifelse(lower_team == WTeamID, WRankWob, LRankWob),
  HTWol = ifelse(higher_team == WTeamID, WRankWol, LRankWol),
  LTWol = ifelse(lower_team == WTeamID, WRankWol, LRankWol),
  HTAvgRank = ifelse(higher_team == WTeamID, WAvgSeed, LAvgSeed),
  LTAvgRank = ifelse(lower_team == WTeamID, WAvgSeed, LAvgSeed),
  DefaultTest = ifelse(WTeamSeed > LTeamSeed, 1, 0)
)

```

Add team names to the detailed data frame for more interpretability
```{r}
# add team names
teams_tourney <- teams %>% select(TeamName, TeamID)
teams_tourney <- teams_tourney %>% rename(lower_team = TeamID)
tourney_d_result <-left_join(x = tourney_d_result,
                                    y = teams_tourney,
                                    by = "lower_team")


teams_tourney <- teams_tourney %>% rename(higher_team = lower_team)
tourney_d_result <- left_join(x = tourney_d_result,
                                     y = teams_tourney, 
                                     by = "higher_team")


# renaming
tourney_d_result %<>% rename(LowerTeamName = TeamName.x,
                                    HigherTeamName = TeamName.y)


```

Add score difference, seed difference and rounds
```{r}
# add score difference
tourney_d_result <- tourney_d_result %>% mutate(scoreDiff = HTScore - LTScore)


# add rounds to tourney_d_results
tourney_d_result <- tourney_d_result %>% mutate(Round = ifelse(
  DayNum %in% 134:135, 0,
  ifelse(DayNum %in% 136:137, 1,
         ifelse(DayNum %in% 138:139, 2,
                ifelse(DayNum %in% 143:144, 3,
                       ifelse(DayNum %in% 145:146, 4,
                              ifelse(DayNum == 152, 5,
                                     ifelse(DayNum == 154, 6, 7))))))
))

tourney_d_result <- tourney_d_result %>% mutate(
  SeedDifference = HTSeed - LTSeed
)

```


Factor some variables
```{r}
# factoring 
tourney_d_result %<>% mutate(Season = factor(Season),
                                    DayNum = factor(DayNum),
                                    HigherTeamName = factor(HigherTeamName),
                                    LowerTeamName = factor(LowerTeamName),
                                    NumOT = factor(NumOT))

tourney_d_result <- tourney_d_result %>% mutate(HTeamConfFactor = fct_lump(HTConf, 10))

tourney_d_result <- tourney_d_result %>% mutate(LTeamConfFactor = 
                                                                fct_lump(LTConf, 10))

```

Make all the in game statistics deltas between the two teams
```{r}
tourney_d_result <- tourney_d_result %>% mutate(
  DFTM = HTFTM - LTFTM,
  DStl = HTStl - LTStl,
  DFGA = HTFGA - LTFGA,
  DFGA3 = HTFGA3 - LTFGA3,
  DFGM3 = HTFGM3 - LTFGM3,
  DFTM = HTFTM - LTFTM,
  DFTA = HTFTA - LTFTA,
  DOR = HTOR - LTOR,
  DDR = HTDR - LTDR,
  DTO = HTTO - LTTO,
  DAst = HTAst - LTAst,
  DBlk = HTBlk - LTBlk,
  DPF = HTPF - LTPF
)

```

Parse the regular season data. Since the regular season tables have slightly different attributes and we won't always use both datasets combined, we will initially parse them separately and then join them as needed.
```{r}
# select all rankings for the last 20 days of the regular season
reg_rankings <- rankings %>% filter(between(rankings$RankingDayNum, 110, 130)) %>% select(Season,
                                                                                      SystemName,
                                                                                      TeamID,
                                                                                      everything())

# get rid of duplicates by system name, season, and team ID
reg_rankings <- reg_rankings[!duplicated(rankings[1:3]),]

# parse rankings by the ranking system
bih_rankings <- rankings %>% filter(SystemName == 'BIH')
col_rankings <- rankings %>% filter(SystemName == 'COL')
dol_rankings <- rankings %>% filter(SystemName == 'DOL')
mor_rankings <- rankings %>% filter(SystemName == 'MOR')
wlk_rankings <- rankings %>% filter(SystemName == 'WLK')
wob_rankings <- rankings %>% filter(SystemName == 'WOB')
wol_rankings <- rankings %>% filter(SystemName == 'WOL')

# rename to prepare for join based
bih_rankings <- bih_rankings %>% rename(WTeamID = TeamID,
                                        WRankBih = OrdinalRank)

col_rankings <- col_rankings %>% rename(WTeamID = TeamID,
                                        WRankCol = OrdinalRank)

dol_rankings <- dol_rankings %>% rename(WTeamID = TeamID,
                                        WRankDol = OrdinalRank)


wlk_rankings <- wlk_rankings %>% rename(WTeamID = TeamID,
                                        WRankWlk = OrdinalRank)

wob_rankings <- wob_rankings %>% rename(WTeamID = TeamID,
                                        WRankWob = OrdinalRank)


wol_rankings <- wol_rankings %>% rename(WTeamID = TeamID,
                                        WRankWol = OrdinalRank)

mor_rankings <- mor_rankings %>% rename(WTeamID = TeamID,
                                        WRankMor = OrdinalRank)

# join with regular season data based on winning team id and season
reg_detailed_result <-left_join(x = regular_detailed_result,
                                    y = bih_rankings,
                                    by = c("Season", "WTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = wob_rankings,
                                    by = c("Season", "WTeamID"))


reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = wol_rankings,
                                    by = c("Season", "WTeamID"))


reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = wlk_rankings,
                                    by = c("Season", "WTeamID"))


reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = dol_rankings,
                                    by = c("Season", "WTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = col_rankings,
                                    by = c("Season", "WTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = mor_rankings,
                                    by = c("Season", "WTeamID"))


# add losing teams rankings to prepare for join
bih_rankings <- bih_rankings %>% rename(LTeamID = WTeamID,
                                        LRankBih = WRankBih)

col_rankings <- col_rankings %>% rename(LTeamID = WTeamID,
                                        LRankCol = WRankCol)

dol_rankings <- dol_rankings %>% rename(LTeamID = WTeamID,
                                        LRankDol = WRankDol)

wlk_rankings <- wlk_rankings %>% rename(LTeamID = WTeamID,
                                        LRankWlk = WRankWlk)

wob_rankings <- wob_rankings %>% rename(LTeamID = WTeamID,
                                        LRankWob = WRankWob)

wol_rankings <- wol_rankings %>% rename(LTeamID = WTeamID,
                                        LRankWol = WRankWol)

mor_rankings <- mor_rankings %>% rename(LTeamID = WTeamID,
                                        LRankMor = WRankMor)


# Running out of memory here - ideally would just configure to this to run on a cloud instance but I will clear some files I don't use anymore
rm(tourney_slots)
rm(tourney_seeds_round_slots)
rm(tourney_games)


# join with regular season data based on losing team id and the season
reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = bih_rankings,
                                    by = c("Season", "LTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = wob_rankings,
                                    by = c("Season", "LTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = wol_rankings,
                                    by = c("Season", "LTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = wlk_rankings,
                                    by = c("Season", "LTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = dol_rankings,
                                    by = c("Season", "LTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = col_rankings,
                                    by = c("Season", "LTeamID"))

reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = mor_rankings,
                                    by = c("Season", "LTeamID"))

# remove unecessary columns
reg_detailed_result = reg_detailed_result[,-grep("^RankingDayNum",colnames(reg_detailed_result))]
reg_detailed_result = reg_detailed_result[,-grep("^SystemName",colnames(reg_detailed_result))]

```


```{r}
# get a certain number of days out of the regular season
reg_detailed_result <- reg_detailed_result %>% filter(between(reg_detailed_result$DayNum, 1, 132))

# get the average seed for each
reg_detailed_result <- reg_detailed_result %>% mutate(
  WAvgSeed = ((WRankMor + WRankWol + WRankWob + WRankWlk + WRankDol + WRankBih
               + WRankCol)/7),
  LAvgSeed = ((LRankMor + LRankWol + LRankWob + LRankWlk + LRankBih + LRankBih
               + LRankCol)/7),
  AvgSeedDiff = WAvgSeed - LAvgSeed
)

```

Add conferences
```{r}
# team conferences to tourney results
team_conferences <- team_conferences %>% rename(WTeamID = TeamID)

# join with conferences
reg_detailed_result <- left_join(x = reg_detailed_result,
                                     y = team_conferences, 
                                     by = c("WTeamID", "Season"))

team_conferences <- team_conferences %>% rename(LTeamID = WTeamID)

reg_detailed_result <- left_join(x = reg_detailed_result,
                                     y = team_conferences, 
                                     by = c("LTeamID", "Season"))

reg_detailed_result <- reg_detailed_result %>% rename(
  WTeamConf = ConfAbbrev.x,
  LTeamConf = ConfAbbrev.y 
)

```

Randomly code data for a binary variable to predict on
```{r}
# now I'm going to look to make a model for lower team wins and higher team wins
reg_detailed_result <- reg_detailed_result %>% mutate(
  lower_team = pmin(WTeamID, LTeamID),
  higher_team = pmax(WTeamID, LTeamID),
  higher_team_won = ifelse(higher_team == WTeamID, 1, 0),
  HTScore = ifelse(higher_team == WTeamID, WScore, LScore),
  LTScore = ifelse(lower_team == WTeamID, WScore, LScore),
  HTConf = ifelse(higher_team == WTeamID, as.character(WTeamConf), as.character(LTeamConf)),
  LTConf = ifelse(lower_team == WTeamID, as.character(WTeamConf), as.character(LTeamConf)),
  HTFGM = ifelse(higher_team == WTeamID, WFGM, LFGM),
  LTFGM = ifelse(lower_team == WTeamID, WFGM, LFGM),
  HTFGA = ifelse(higher_team == WTeamID, WFGA, LFGA),
  LTFGA = ifelse(lower_team == WTeamID, WFGA, LFGA),
  HTFGM3 = ifelse(higher_team == WTeamID, WFGM3, LFGM3),
  LTFGM3 = ifelse(lower_team == WTeamID, WFGM3, LFGM3),
  HTFGA3 = ifelse(higher_team == WTeamID, WFGA3, LFGA3),
  LTFGA3 = ifelse(lower_team == WTeamID, WFGA3, LFGA3),
  HTFTM = ifelse(higher_team == WTeamID, WFTM, LFTM),
  LTFTM = ifelse(lower_team == WTeamID, WFTM, LFTM),
  HTFTA = ifelse(higher_team == WTeamID, WFTA, LFTA),
  LTFTA = ifelse(lower_team == WTeamID, WFTA, LFTA),
  HTOR = ifelse(higher_team == WTeamID, WOR, LOR),
  LTOR = ifelse(lower_team == WTeamID, WOR, LOR),
  HTDR = ifelse(higher_team == WTeamID, WDR, LDR),
  LTDR = ifelse(lower_team == WTeamID, WDR, LDR),
  HTAst = ifelse(higher_team == WTeamID, WAst, LAst),
  LTAst = ifelse(lower_team == WTeamID, WAst, LAst),
  HTTO = ifelse(higher_team == WTeamID, WTO, LTO),
  LTTO = ifelse(lower_team == WTeamID, WTO, LTO),
  HTStl = ifelse(higher_team == WTeamID, WStl, LStl),
  LTStl = ifelse(lower_team == WTeamID, WStl, LStl),
  HTBlk = ifelse(higher_team == WTeamID, WBlk, LBlk),
  LTBlk = ifelse(lower_team == WTeamID, WBlk, LBlk),
  HTPF = ifelse(higher_team == WTeamID, WPF, LPF),
  LTPF = ifelse(lower_team == WTeamID, WPF, LPF),
  HTBih = ifelse(higher_team == WTeamID, WRankBih, LRankBih),
  LTBih = ifelse(lower_team == WTeamID, WRankBih, LRankBih),
  HTCol = ifelse(higher_team == WTeamID, WRankCol, LRankCol),
  LTCol = ifelse(lower_team == WTeamID, WRankCol, LRankCol),
  HTDol = ifelse(higher_team == WTeamID, WRankDol, LRankDol),
  LTDol = ifelse(lower_team == WTeamID, WRankDol, LRankDol),
  HTMor = ifelse(higher_team == WTeamID, WRankMor, LRankMor),
  LTMor = ifelse(lower_team == WTeamID, WRankMor, LRankMor),
  HTWlk = ifelse(higher_team == WTeamID, WRankWlk, LRankWlk),
  LTWlk = ifelse(lower_team == WTeamID, WRankWlk, LRankWlk),
  HTWol = ifelse(higher_team == WTeamID, WRankWol, LRankWol),
  LTWol = ifelse(lower_team == WTeamID, WRankWol, LRankWol)
)

```

Add team names for interpretability and add score difference
```{r}
# add team names
teams <- read.csv("Teams.csv")
teams <- teams %>% select(TeamName, TeamID)
teams <- teams %>% rename(lower_team = TeamID)
reg_detailed_result <-left_join(x = reg_detailed_result,
                                    y = teams,
                                    by = "lower_team")


teams <- teams %>% rename(higher_team = lower_team)
reg_detailed_result <- left_join(x = reg_detailed_result,
                                     y = teams, 
                                     by = "higher_team")


# renaming
reg_detailed_result %<>% rename(LowerTeamName = TeamName.x,
                                    HigherTeamName = TeamName.y)

# add score difference
reg_detailed_result <- reg_detailed_result %>% mutate(scoreDiff = HTScore - LTScore)
```

Factor a variety of variables
```{r}
# factoring 
reg_detailed_result %<>% mutate(Season = factor(Season),
                                    DayNum = factor(DayNum),
                                    HigherTeamName = factor(HigherTeamName),
                                    LowerTeamName = factor(LowerTeamName),
                                    NumOT = factor(NumOT))


reg_detailed_result <- reg_detailed_result %>% mutate(HTeamConfFactor = fct_lump(HTConf, 10))

reg_detailed_result <- reg_detailed_result %>% mutate(LTeamConfFactor = 
                                                                fct_lump(LTConf, 10))

```


Add deltas for the in game statistics
```{r}
# transition all wins and losses into deltas
reg_detailed_result <- reg_detailed_result %>% mutate(
  DFTM = HTFTM - LTFTM,
  DFGM = HTFGM - LTFGM,
  DStl = HTStl - LTStl,
  DFGA = HTFGA - LTFGA,
  DFGA3 = HTFGA3 - LTFGA3,
  DFGM3 = HTFGM3 - LTFGM3,
  DFTM = HTFTM - LTFTM,
  DFTA = HTFTA - LTFTA,
  DOR = HTOR - LTOR,
  DDR = HTDR - LTDR,
  DTO = HTTO - LTTO,
  DAst = HTAst - LTAst,
  DBlk = HTBlk - LTBlk,
  DPF = HTPF - LTPF
)
```

Get averages
```{r}
averages <- reg_detailed_result %>% 
  group_by(higher_team, Season) %>% 
  summarize(DFTMReg = mean(DFTM, na.rm = TRUE),
            DFGMReg = mean(DFGM, na.rm = TRUE),
            DStlReg = mean(DStl, na.rm = TRUE),
            DFGAReg = mean(DFGA, na.rm = TRUE),
            DFGA3Reg = mean(DFGA3, na.rm = TRUE),
            DFGM3Reg = mean(DFGM3, na.rm = TRUE),
            DFTAReg = mean(DFTA, na.rm = TRUE),
            DORReg = mean(DOR, na.rm = TRUE),
            DDRReg = mean(DDR, na.rm = TRUE),
            DTOReg = mean(DTO, na.rm = TRUE),
            DAstReg = mean(DAst, na.rm =TRUE),
            DBlkReg = mean(DBlk, na.rm = TRUE),
            DPFAReg = mean(DPF, na.rm = TRUE))

```


Combine with tourney detailed result
```{r}


```


Combine with views and ratings data
```{r}
# factor the season
rating_round_viewers <- rating_round_viewers %>% mutate(Season = factor(Season))
tourney_semi_views_ratings <- tourney_semi_views_ratings %>% mutate(
  Season = factor(Season)
)

# join the ratings with the tourney detailed result
tourney_finals_views_ratings <-left_join(x = tourney_d_result,
                                    y = rating_round_viewers,
                                    by = c("Season", "Round"))

tourney_semi_views_ratings <- tourney_semi_views_ratings %>% rename(WTeamID = TeamID1)
tourney_semi_views_ratings <- tourney_semi_views_ratings %>% rename(LTeamID = TeamID2)

# data frame with the semi final ratings and viewers
tourney_semi_views_ratings <- left_join(x = tourney_detailed_result,
                                        y = tourney_semi_views_ratings,
                                        by= c("Season", "WTeamID", "LTeamID"))

# data frame with the finals ratings and viewers
tourney_finals_views_ratings <- tourney_finals_views_ratings %>% filter(
  Round == 6
)

tourney_semi_views_ratings <- tourney_semi_views_ratings %>% rename(Round = Round.y)
tourney_semi_views_ratings <- tourney_semi_views_ratings %>% select(-Round.x)

tourney_views_ratings <- rbind(tourney_semi_views_ratings, tourney_finals_views_ratings)

tourney_views_ratings <- tourney_views_ratings %>% mutate(
  Network = factor(Network)
)

```

Add some variables that might be interesting to analyze in relation to the views and ratings
```{r}
tourney_views_ratings <- tourney_views_ratings %>% mutate(
  IsUpset = ifelse(WTeamSeed < LTeamSeed, 1, 0)
)

tourney_views_ratings <- tourney_views_ratings %>% mutate(
  SeedDifference = abs(WTeamSeed - LTeamSeed)
)

tourney_views_ratings <- tourney_views_ratings %>% mutate(
  TotalThreesMade = WFGM3 + LFGM3,
  TotalFieldGoals = WFGM + LFGM,
  TotalFieldGoalsAttempts = WFGA + LFGA,
  TotalPointsScored = WScore + LScore
)


```



